---
title: "EP Political Groups' Cohesion Rates"
author:
  - Marco SCIPIONI
  - Filipa TORRAO
date: "`r Sys.Date()`"
format: 
 docx:
   toc: true
   toc-depth: 3
   toc-title: Contents
   number-sections: true
   colorlinks: true
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  warning: false
---

## KEY MESSAGES

* The **Greens** display the highest **vote participation** among all Groups.
*__Renew__ comes third*. 
* **S&D** is the most cohesive political group, with **Renew** being the *third* most cohesive.
* Political Groups on the **extremes** of the political spectrum tend to be *less cohesive on final votes*, with the exception of ESN.
Conversely, Groups in the **centre** tend to be *more cohesive on final votes*, with the exception of Renew.
* Renew appears as the most cohesive in CULT, INTA, ITRE, and PECH. Worst cohesion in EMPL.
* The LEFT, ESN, and PfE have largely increased their cohesion rate across the mandate, with centre parties remaining stable.

## Intro
```{r}
#| include: false

script_starts <- Sys.time()

###--------------------------------------------------------------------------###
## Libraries -------------------------------------------------------------------
if ( !require("pacman") ) install.packages("pacman")
pacman::p_load(char = c("data.table", "dplyr", "forcats", "ggplot2", "here",
                        "lubridate", "janitor", "stringr", "tidyr", "tidyselect",
                        "tidytext") )


# Hard code the start of the mandate ------------------------------------------#
mandate_starts <- "2024-07-14"


###--------------------------------------------------------------------------###
## Functions -------------------------------------------------------------------

# Mode ------------------------------------------------------------------------#
# https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode
stat_mode <- function(x) {
  if ( length(x) <= 2 ) return(x[1])
  if ( anyNA(x) ) x = x[!is.na(x)]
  ux <- unique(x)
  ux[ which.max( tabulate( match(x, ux) ) ) ] }

# Load join functions ---------------------------------------------------------#
source(file = here::here("scripts_r", "join_functions.R") )

# Calculate Majorities --------------------------------------------------------#
source(file = here::here("scripts_r", "get_majority.R") )


#------------------------------------------------------------------------------#
## Graphics --------------------------------------------------------------------
# set theme globally for ggplots ----------------------------------------------#
# https://stackoverflow.com/questions/34522732/changing-fonts-in-ggplot2
theme_set(theme_minimal(base_size = 10, base_family = "roboto") )

#------------------------------------------------------------------------------#
# vote colours ----------------------------------------------------------------#
vote_colours <- c(For = '#00AEEF',
                  Against = '#BE3455',
                  Abstain = "#969696",
                  `Did not vote` = '#5D5CA4',
                  Absent = '#F47920')

# political groups colours ----------------------------------------------------#
polgroup_cols = c(`S&D` = "#EE3652",
                  ECR = "#0D88C3",
                  PfE = "#1A3153",
                  PPE = "#3C5979",
                  Renew = "#FFCC70",
                  `Verts/ALE` = "#19A24A",
                  `The Left` = "#733542",
                  ESN = "#000000",
                  NI = "#979797")
```

The objective of this short note is to investigate trends in **cohesion rates** during the EP's current mandate.

**Please keep in mind** that this document is the *sole property of the Renew Group* and *should not be disseminated to other people or organisations*.


## Data and Methods
```{r}
#| include: false

###--------------------------------------------------------------------------###
## Read data -------------------------------------------------------------------
# RCV
meps_rcv_mandate <- data.table::fread(
  file = here::here("data_out", "meps_rcv_mandate_10.csv"),
  verbose = TRUE, key = c("rcv_id", "pers_id"), na.strings = c(NA_character_, "") )

# Recode
vote_dict <- data.table::fread(here::here("data_out", "votes", "vote_dict.csv") )
# left join
# dim(meps_rcv_mandate)
meps_rcv_mandate <- vote_dict[meps_rcv_mandate, on = "result"]
# dim(meps_rcv_mandate)
meps_rcv_mandate[, result_fct := as.factor(result_fct)]

# VOTES
pl_votes <- data.table::fread(
  file = here::here("data_out", "votes", "pl_votes_10.csv"), 
  # select = c("activity_date", "rcv_id", "mandate", "number_of_attendees", 
  #            "number_of_votes_abstention", "number_of_votes_against",
  #            "number_of_votes_favor"),
  na.strings = c(NA_character_, "") 
)

meps_current = data.table::fread(here::here(
  "data_out", "meps", "meps_current.csv") )

### Committees ----------------------------------------------------------------#
# List of RCV IDs and Committees
rcvid_cmts <- data.table::fread(here::here(
  "data_out", "rcv", "rcvid_cmts_10.csv") )
rcvid_cmts[committee_lab == "BUDE", committee_lab := "BUDG"]

# Number of RCVs
# length( unique(meps_rcv_mandate$rcv_id) )
# Number of RCVs by Committee (check for Joint Committee)
# rcvid_cmts[, .N, by = rcv_id][order(N)]

# what are the RCVs without DOC ID?
rcv_missing_doc <- unique(meps_rcv_mandate$rcv_id)[
  !unique(meps_rcv_mandate$rcv_id) %in% unique(rcvid_cmts$rcv_id)]

# These should be ordre du jour, election of commission, etc.
# pl_votes[rcv_id %in% rcv_missing_doc]

#------------------------------------------------------------------------------#
# get length objects
n_votes <- nrow(pl_votes)
n_rcv <- length( unique( meps_rcv_mandate$rcv_id ) )


#------------------------------------------------------------------------------#
# Parameters -------------------------------------------------------------------
renew_polgroup_id <- 7035L

renew_current_meps = unique(meps_current$pers_id[
  meps_current$polgroup_id == renew_polgroup_id
])
```

We include in our analysis all voting data relative to roll-call votes (RCVs) since `r mandate_starts`.
To provide some context, we currently have `r length(unique(meps_rcv_mandate$rcv_id))` RCVs in our database.
It is important to appreciate that what we have is only a subset of the all legislation being adopted. 
Indeed, the EP recognises different forms of voting - such as *raise of hands*, or *electronic votes* - but RCVs are the only ones where we able to trace votes to individual MEPs - that is, they are so-called *nominal* votes. 
Since the start of the current mandate, RCVs were approximately `r round(n_rcv/n_votes*100, digits=1)`% of the total votes.


## Participation in the votes
We define vote participation as the share of total RCVs wherein MEPs did vote.
To be clear, we do not take into account the Attendance Registers.
Thus, if a MEP was officially '*excused*', we would lose that. 
That being said, if that is of interest, we could always include those cases. 

The **Greens** display the highest **vote participation** among all Groups.
```{r}
#| fig-width: 8
#| fig-height: 6

#------------------------------------------------------------------------------#
meps_rcv_mandate[, `:=`(
  has_voted = data.table::fifelse(
    test = result >= -1L, yes = 1L, no = 0L) 
) ]

voted_dt <- meps_rcv_mandate[
  !is.na(has_voted)
  & !is.na(polgroup_id), 
  .N, 
  keyby = list(polgroup_id, pers_id, has_voted)]    
voted_dt[, tot := sum(N, na.rm = TRUE), by = list(polgroup_id, pers_id)]
voted_dt[, attendance_rate := round(N / tot, digits = 3)]
attendance_bypolgroup = voted_dt[
  has_voted == 1L, 
  list(attendance_rate_avg = mean(attendance_rate, na.rm=TRUE)),
  by = list(polgroup_id)
] |> 
  join_polit_labs() 


# Plot ------------------------------------------------------------------------#
# REF: https://stackoverflow.com/questions/11020437/consistent-width-for-geom-bar-in-the-event-of-missing-data?noredirect=1&lq=1
attendance_bypolgroup |> 
  dplyr::mutate(political_group = forcats::fct_reorder(.f = political_group, 
                                                       .x = attendance_rate_avg, 
                                                       .desc = TRUE)) |>
  ggplot(aes(x = attendance_rate_avg, y = fct_rev(political_group),
             fill = political_group) ) +
  geom_col(linewidth=0.1, colour = "black", show.legend = FALSE) +
  geom_text(aes(label = round(attendance_rate_avg*100, digits = 1)), 
            hjust = -0.5, size = 4) + 
  scale_fill_manual(
    values = polgroup_cols
  ) +
  labs(
    x = "Participation Rate (%)", y = "", fill = "",
    caption = stringr::str_wrap("Source: EP. Notes: data is relative to RCVs in Plenary. The height of the bars represent the mean of the distribution of the attandance rates during RCVs. Attendance rate is defined as participation in RCVs.",
                                width = 150) ) +
  scale_x_continuous(limits=c(0.5, 1), 
                     oob=scales::rescale_none,
                     labels = scales::percent) +
  theme(
    axis.text.y = element_text(face = "bold", size = 12)
  )
```


## Cohesion Rates
The formula for cohesion is taken from [Is there a selection bias in roll call votes?](http://link.springer.com/10.1007/s11127-018-0529-1)

$COHESION_{ij} = \frac{max[Y_{ij}, N_{ij}, A_{ij}] - 0.5[(Y_{ij} + N_{ij} + A_{ij}) - max[Y_{ij}, N_{ij}, A_{ij}]]}{(Y_{ij} + N_{ij} + A_{ij})}$

To develop an intuition for how this metric works, please consider the 6 simplified scenarios in the table below.
In each row, a party of different size (namely, 2, 3, and 4 MEPs) has a different distribution of votes internally. 
The rightmost column provides the cohesion rate for each of these cases.
```{r}
knitr::kable(
  tibble::tibble(
    `Number of MEPs` = c(2, 2, 3, 3, 4, 4),
    `Vote: For` = c(2, 1, 2, 1, 2, 3),
    `Vote: Against` = c("-", 1, 1, 1, 1, 1),
    `Vote: Abstain` = c("-", "-", "-", 1, 1, "-"), 
    Cohesion = c(100, 25, 50, 0, 25, 62.5),
    # `Is split?` = c("No", "Yes", "No", "Yes", "Yes", "No")
  ), 
  align = "c")
```

As a *technical side note*, please be aware that this basically picks the mode of the 3 typologies of vote officially recognised in the EP. 
However, experts have highlighted other forms of vote are actually at the MEPs' disposal, i.e. being present in the House but not showing up for specific votes (`no_vote`), or being absent on the voting day (`absent`).
We do not count those in this note, but can be included if needs be.
Further, and to be clear, here we do not make use of data coming from the Attendance Registers, but these can be included if needs be.

### Cohesion Rates in Plenary
#### Overall cohesion rates
```{r}
#| fig-width: 8
#| fig-height: 6

###--------------------------------------------------------------------------###
# Read in function
source(file = here::here("scripts_r", "cohesionrate_function.R") )
# convert to factor - essential for tabulate
meps_rcv_mandate[, result_fct := factor(result,
                                        levels = -3L : 1L,
                                        labels = c("absent", "no_vote", "against",
                                                   "abstain", "for") ) ]
# Check
# table(meps_rcv_mandate$result, meps_rcv_mandate$result_fct, exclude = NULL)

# Calculate Cohesion
cohesion_dt <- meps_rcv_mandate[
  result >= -1L, # only official votes
  list(
    cohesion = cohesion_hn(result_fct),
    max_mode = max( tabulate(result_fct) ),
    tot_votes = sum( tabulate(result_fct) ) ), 
  keyby = list(rcv_id, polgroup_id) ] |> 
  join_polit_labs()

cohesion_dt[, .(avg = round( mean(cohesion, na.rm = TRUE), digits = 1)), 
            by = list(political_group)][order(-avg)] |> 
  dplyr::mutate(political_group = fct_reorder(.f = political_group, .x = avg)) |> 
  ggplot(aes(x=political_group, y = avg, fill = political_group)) +
  geom_col(colour = "black", linewidth = 0.1) +
  geom_text(aes(label = avg), size = 3, nudge_y = 3) +
  coord_flip() +
  labs(x="", y="Average Cohesion Rates (%)", fill="",
       title = "Cohesion Rates by EP Political Groups",
       caption = "Source: EP. Notes: data is relative to RCVs in Plenary") +  
  scale_fill_manual(values = polgroup_cols) +
  guides(fill = guide_legend(nrow = 1)) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "none",
        # strip.text.y.right = element_text(angle = 0, face = "bold"),
        axis.text.y = element_text(face = "bold"))
# ggsave(filename = here::here("figures", "cohesion_bypolgroup_avg.pdf"), 
#        device = "pdf", dpi = 300, height = 8, width = 12)
```


*S&D are the most cohesive group to date (93,3% cohesion index), during the current mandate, followed by the Greens (92,8%), and finally Renew (91,6%). Patriots for Europe are the least cohesive group. Overall, we can observe that the centre parties are in general more cohesive than the extremes counterparts of the political spectrum.*


#### Cohesion Rates by Final
```{r}
#| fig-width: 8
#| fig-height: 5

#------------------------------------------------------------------------------#
# Final Votes data ------------------------------------------------------------#
votes_final_all <- data.table::fread( here::here(
  "data_out", "votes", "votes_final_10.csv") )

# Merge RCV with Finals -------------------------------------------------------#
meps_rcv_mandate <- votes_final_all[meps_rcv_mandate, on = "rcv_id"]

# Colours ---------------------------------------------------------------------#
cols_final <- c(Final = "#0868ac", `Not Final` = "#d7301f")

# Calculate final votes -------------------------------------------------------#
cohesion_final_dt <- meps_rcv_mandate[
  result >= -1L, # only official votes
  list(
    cohesion = cohesion_hn(result_fct),
    max_mode = max(tabulate(result_fct)),
    tot_votes = sum(tabulate(result_fct) ) ),
  keyby = list(rcv_id, polgroup_id, is_final) ] |> 
  join_polit_labs()

# Plot -------------------------------------------------------------------------
# REF: https://stackoverflow.com/questions/40211451/geom-text-how-to-position-the-text-on-bar-as-i-want
cohesion_final_avg <- cohesion_final_dt[, list(
  avg = round( mean(cohesion, na.rm = TRUE), digits = 1)), 
  by = list(political_group, is_final)][order(avg)]

# Extract rank vector for plot
plot_rank <- cohesion_final_avg$political_group[
  cohesion_final_avg$is_final == 1L]

cohesion_final_avg |> 
  dplyr::mutate(
    is_final = ifelse(test = is_final == 1L,
                      yes = "Final", no = "Not Final"), 
    is_final = factor(is_final, levels = c("Final", "Not Final") ) ) |> 
  ggplot(aes(x = factor(political_group, levels = plot_rank), y = avg) ) +
  geom_col(aes(fill = is_final, group = is_final),
           colour = "black", linewidth = 0.1, position = "dodge") +
  geom_text(aes(label = avg, group = is_final), size = 3,
            hjust = -0.1, position = position_dodge(width = 1), inherit.aes = TRUE) +
  scale_fill_manual(values = cols_final) +
  coord_flip() +
  labs(x="", y="Average Cohesion Rates (%)", fill="",
       title = "Cohesion Rates by EP Political Groups",
       caption = "Source: EP. Notes: data is relative to RCVs in Plenary") +  
  guides(fill = guide_legend(nrow = 1)) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "top",
        # strip.text.y.right = element_text(angle = 0, face = "bold"),
        axis.text.y = element_text(face = "bold") )
# ggsave(filename = here::here("figures", "cohesion_bypolgroup_avg.pdf"), 
#        device = "pdf", dpi = 300, height = 8, width = 12)
```


*In terms of the final votes, we further observe an increase in the cohesion of the centre parties, now only separated by 3,7% (PPE vs S&D). Extremes of the political spectrum tend to be less cohesive on final votes, than on not final votes (with the exception of ESN). On the contrary, Centre parties are more cohesive on final than on non final votes (Renew being the exception, but only marginally by 0,6%).*


#### Cohesion Rates by Committee
```{r}
#| fig-width: 8
#| fig-height: 13

#------------------------------------------------------------------------------#
cohesion_bycommittee <- rcvid_cmts[
  cohesion_dt,
  on = "rcv_id", nomatch = NULL
] 

cohesion_bycommittee[
  !is.na(committee_lab),
  list(avg_cohesion = mean(cohesion) ),
  by = list(political_group, committee_lab)] |> 
  dplyr::mutate(political_group_rw = reorder_within(
    political_group, avg_cohesion, committee_lab) ) |>
  ggplot(aes(x=political_group_rw, y = avg_cohesion, fill = political_group)) +
  geom_col(colour = "black", linewidth = 0.1) +
  facet_wrap(~committee_lab, scales = "free_y", ncol = 2) +
  coord_flip() +
  labs(x="", y="Average Cohesion Rates (%)", fill="",
       title = "Cohesion Rates by Committees",
       caption = "Source: EP. Notes: data is relative to RCVs in Plenary") +  
  scale_x_reordered() +
  scale_fill_manual(values = polgroup_cols) +
  guides(fill = guide_legend(nrow = 1)) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "none",
        # strip.text.y.right = element_text(angle = 0, face = "bold"),
        axis.text.y = element_text(face = "bold"))
# ggsave(filename = here::here("figures", "cohesion_bycommittee.pdf"),
#        device = "pdf", dpi = 300, height = 8, width = 12)
```


*When we take a look at committees, Renew appears as the most cohesive in CULT, INTA, ITRE, JURI (tied), and PECH, with a worseperformance (less cohesive) in SANT, REGI, and especially in EMPL, where it appears only above the NI members. EPP only tops the cohesion charts in ECON, closely followed by Renew. *

#### Cumulative Cohesion Rates
```{r}
#| fig-width: 8
#| fig-height: 9.5

#------------------------------------------------------------------------------#
rcvid_date <- unique(
  pl_votes[, list(date = activity_date, rcv_id)]
)
last_pl_month = lubridate::month( max(pl_votes$activity_date), 
                                  label = TRUE, abbr = TRUE)
last_pl_year = data.table::year( max(pl_votes$activity_date) )
month_year = paste(last_pl_month, last_pl_year, sep = " ")


# All Political Groups --------------------------------------------------------#
cohesion_allpolgroups <- rcvid_date[
  meps_rcv_mandate, 
  on = "rcv_id", nomatch = NULL
][result >= -1L, # only official votes
  list(cohesion_rate = cohesion_hn(result_fct)),
  keyby = list(date, rcv_id, polgroup_id)
][!is.na(cohesion_rate), 
  `:=`(
    cum_cohesion = dplyr::cummean(cohesion_rate)
  ),
  by = list(polgroup_id)
] |> 
  join_polit_labs()

cohesion_allpolgroups[
  political_group != "NI"
][
  order(date, rcv_id)
][
  , `:=`(
    rcv_id = as.factor(rcv_id),
    political_group = factor(political_group, 
                             levels = c("Renew", "PPE", "S&D", "Verts/ALE",
                                        "ECR", "The Left", "PfE", "ESN") ) 
  )
] |>
  ggplot(aes(x = rcv_id, y = cum_cohesion, group = political_group,
             colour = political_group, fill = political_group) ) +
  geom_point(alpha = 0.1, show.legend = FALSE) +
  geom_smooth(show.legend = FALSE) +
  facet_wrap(~political_group, nrow=4) + # change this to 2 if printing pdf
  scale_x_discrete(
    breaks = c(as.character(min(cohesion_allpolgroups$rcv_id)),
               as.character(max(cohesion_allpolgroups$rcv_id))),
    labels = c("7/2024", month_year) ) +
  labs(x="Time", y="Cumulative Average Cohesion Rate") +
  scale_colour_manual(values = polgroup_cols) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust=c(0,1)),
    panel.spacing.x = unit(5, "mm")
  )
# ggsave(filename = here::here("figures", "cumulative_cohesion_allgroups.pdf"),
#        device = "pdf", dpi = 300, height = 7, width = 12.5)
```

*For the cumulative cohesion rates, the centre parties tend to have a stable cohesion, with S&D and the Greens showing a very slight increase over time. The LEFT, ESN, and PfE have largely increased their cohesion rate across the mandate.*


### Cohesion Rates in Committees
```{r}
#| fig-width: 8
#| fig-height: 9

###--------------------------------------------------------------------------###
# !!WATCH OUT!! ---------------------------------------------------------------#
# Grab data from Committee repo - CHECK YOUR PATHS!
###--------------------------------------------------------------------------###

cohesion_aggr <- data.table::fread(
  file = "/home/scipima/Documents/github/iter_committee_test/data_out/tmp/cohesion_aggr.csv")
cohesion_cmt <- data.table::fread(
  file = "/home/scipima/Documents/github/iter_committee_test/data_out/tmp/cohesion_cmt.csv")


cohesion_cmt |> 
  mutate(political_group = forcats::fct_reorder(.f = political_group, 
                                                .x = cohesion, 
                                                .fun = mean)) |> 
  ggplot(aes(x = cohesion, y = political_group)) +
  geom_jitter(aes(alpha = tot_votes, size = tot_votes), 
              shape = 21, width = 0, height = 0.3, linewidth=0.1,
              colour = "black", fill = "grey") +
  stat_summary(aes(fill = political_group),
               fun=mean, geom="point", size = 10, shape = 23, linewidth=2, 
               show.legend = FALSE) +
  scale_fill_manual(values = polgroup_cols) +
  labs(x="Cohesion Rates (%)", y = "" , 
       size="N. MEPs by Vote", alpha = "N. MEPs by Vote",
       title = "EPP's Cohesion Rates by Committee and Vote Type",
       caption = "Source: EP. Notes: data is relative to RCVs in Plenary") +  
  guides(fill = guide_legend(nrow = 1)) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 18),
        plot.subtitle = element_text(size = 15),
        legend.position = "top",
        # strip.text.y.right = element_text(angle = 0, face = "bold"),
        axis.text.y = element_text(face = "bold"))
ggsave(filename = here::here("figures", "cohesion_cmt.jpeg"),
       device = "jpeg", dpi = 300, height = 9, width = 8)
```

S&D is the most cohesive Political Group emerging from Committee RCVs.
Renew comes fourth, after the Greens and ESN. 
However, it must be noted that ESN's results are likely skewed because of the extremely large share of votes where there was just 1 MEP attending the votes, and thus calculating the metric is unfeasible. 
```{r}
cohesion_aggr |> 
  dplyr::arrange(desc(metric_missing)) |> 
  knitr::kable(
    col.names = c("Political Group", "No metric", "Metric present", "Total"), 
    align = "c")

```


## Reproducibility
```{r reproducibility}
sessionInfo()

script_ends <- Sys.time()
print(script_ends - script_starts) # running time
```
