---
title: "Targeting Renew Europe-compatible MEPs outside the Group"
author:
  - Marco SCIPIONI
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    toc-title: Contents
    number-sections: true
    colorlinks: true
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  warning: false
---

## Intro
```{r}
#| include: false

#------------------------------------------------------------------------------#
## Libraries -------------------------------------------------------------------
if ( !require("pacman") ) install.packages("pacman")
pacman::p_load(char = c("data.table", "dplyr", "forcats", "ggplot2", "here",
                        "lubridate", "janitor", "readr", "tidyr", "tidyselect",
                        "tidytext") )


#------------------------------------------------------------------------------#
# Hard code the start of the mandate ------------------------------------------#
mandate_starts <- as.Date("2019-07-01")


#------------------------------------------------------------------------------#
## Functions -------------------------------------------------------------------
# https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode
stat_mode <- function(x) {
  if ( length(x) <= 2 ) return(x[1])
  if ( anyNA(x) ) x = x[!is.na(x)]
  ux <- unique(x)
  ux[ which.max( tabulate( match(x, ux) ) ) ] }

# Load join functions ---------------------------------------------------------#
source(file = here::here("scripts_r", "join_functions.R") )  

# Calculate Majorities --------------------------------------------------------#
source(file = here::here("scripts_r", "get_majority.R") )


###--------------------------------------------------------------------------###
## Graphics --------------------------------------------------------------------
# Set global theme for ggplot2 ------------------------------------------------#
# https://stackoverflow.com/questions/34522732/changing-fonts-in-ggplot2
theme_set(theme_minimal(base_size = 10))

# vote colours  -------------------------------------------------------------###
vote_colours <- c(For = '#00AEEF',
                  Against = '#BE3455',
                  Abstain = "#969696",
                  `Did not vote` = '#5D5CA4',
                  Absent = '#F47920')

# political groups colours ----------------------------------------------------#
polgroup_cols = c(`S&D` = "#EE3652",
                  ECR = "#0D88C3",
                  PfE = "#1A3153",
                  PPE = "#3C5979",
                  Renew = "#FFCC70",
                  `Verts/ALE` = "#19A24A",
                  `The Left` = "#733542",
                  ESN = "#000000",
                  NI = "#979797")
```

```{r}
#| include: false

###--------------------------------------------------------------------------###
## Read data -------------------------------------------------------------------
### RCV ------------------------------------------------------------------------
meps_rcv_mandate <- data.table::fread(
  file = here::here("data_out", "meps_rcv_mandate_all.csv"),
  verbose = TRUE, key = c("rcv_id", "pers_id"), 
  na.strings = c(NA_character_, "") )
# Nominal change from GUE to The Left
meps_rcv_mandate[polgroup_id == 5151L, polgroup_id := 6259L] 

### Votes ----------------------------------------------------------------------
votes_dt <- data.table::fread(
  file = here::here("data_out", "votes", "votes_dt_all.csv"), 
  # select = c("activity_date", "rcv_id", "mandate", "number_of_attendees", 
  #            "number_of_votes_abstention", "number_of_votes_against",
  #            "number_of_votes_favor"), 
  na.strings = c(NA_character_, "") )


### MEP last Plenary day -------------------------------------------------------
meps_current <- data.table::fread( file = here::here(
  "data_out", "meps", "meps_current.csv") )


#------------------------------------------------------------------------------#
# get length objects
n_votes <- nrow(votes_dt)
n_rcv <- length( unique( meps_rcv_mandate$rcv_id ) )

#------------------------------------------------------------------------------#
# Hard code Groups
renew_group_ids <- political_groups$identifier[
  political_groups$label == "Renew"]

# Hard code target MEPs 
target_meps <- c(197699L, 256952L)
target_meps_polgroups <- c(7036L, 6259L, 5151L)
```

The objectives of this short note is to identify Renew Europe-compatible MEPs outside the group. 
It does that by calculating the  **affinity** (also called **similarity**) between Renew and other MEPs.

**Please keep in mind** that this document is the *sole property of the Renew Group* and *should not be disseminated to other people or organisations*.


## Affinity between Renew and other Groups' MEPs
In this section we calculate the affinity of two MEPs currently outside Renew with Renew's majority.
In all visualisations below, we display also the top 5 MEPs according to the same metrics from EPP and S&D who have been *members during both the 9th and 10th mandates*, so that the data become from informative. 
We do this in two different ways, namely by calculating both the **Euclidean distance** and the share of exact matches in votes.
The former metric is more granular, enabling us to capture more profound disagreements between majorities and/or MEPs, whereas the latter only check whether the vote was identical.

In the first subsection below we display the overall affinity, whereas in the second subsection we break it down by Committees.


### Overall affinity

::: {.panel-tabset}

#### Euclidean distance
```{r}
# Get the Groups' majority by rcv_id ------------------------------------------#
mjrt_polgroup <- get_polgroup_majority(
  data_in = meps_rcv_mandate[
    result >= -1] # exclude absent & no votes
)

# Get the target party majority
renew_mjrt <- mjrt_polgroup[
  polgroup_id %in% renew_group_ids]

# Inner-join
renew_othermeps <- renew_mjrt[
  meps_rcv_mandate[
    (pers_id %in% target_meps
     | polgroup_id %in% c(5153L, 7018L, # EPP
                          5154L, 7038L) ) #S&D
    & result >= -1L],
  on = "rcv_id", nomatch = NULL]

# Calculate Euclidean distance
renew_othermeps[, `:=`(
  is_same = as.integer(result == i.result),
  eucl_dist = sqrt((result - i.result)^2) # 0 if same, 2 if distant
)]

# take the average for plotting
dist_avg_dt <- renew_othermeps[, list(
  eucl_dist_avg = mean(eucl_dist, na.rm = TRUE) ),
  by = list(pers_id)
][, `:=`(
  # normalise the euclidean for comparable plotting
  eucl_dist_avg_norm = eucl_dist_avg / 2L
)]
```

Affinity may be defined as the [Euclidean distance](https://en.wikipedia.org/wiki/Euclidean_distance) between Renew's majority and all other MEPs.
The assumption here is that there is a qualitative difference between, say, *abstaining* and voting *against*, with the latter expressing a more profound disagreement with those who voted in *favour*.
To measure affinity, we consider only EP official votes, that is `for`, `against`, `abstain`.

When reading these figures, remember that **low numbers stand for high affinity**, whereas **high numbers represent low affinity**.
The metric is bounded 0-1.
```{r}
#| fig-width: 8
#| fig-height: 4

# plot -------------------------------------------------------------------------
dist_avg_dt |>  
  dplyr::inner_join(
    y = meps_current[
      polgroup_id %in% c(7018L, 7038L)
      | pers_id %in% target_meps
      | polgroup_id %in% target_meps_polgroups],
    by = "pers_id") |>
  join_polit_labs() |>
  join_meps_names() |>
  dplyr::slice_min(order_by = eucl_dist_avg_norm, n = 5, with_ties = FALSE, by = political_group) |> 
  dplyr::mutate(mep_name = tidytext::reorder_within(
    x = mep_name, by = eucl_dist_avg_norm, within = political_group) ) |> 
  # dplyr::filter(political_group == "PPE") |>
  ggplot(aes(x = mep_name, 
             y = eucl_dist_avg_norm, fill = political_group) ) +
  geom_col(colour="black", linewidth=0.1, show.legend = FALSE) +
  geom_text(aes(label = round(eucl_dist_avg_norm, digits = 4)), 
            nudge_y = .01, size = 5) +
  geom_hline(yintercept = 0, linewidth=0.5) +
  facet_grid(cols = vars(political_group), scales = "free_x", space = "free") +
  scale_x_reordered() +
  ylim(0, 1) +
  labs(
    x="", y="Normalised euclidean distance", 
    title = "Distance between Renew's majority and other MEPs",
    subtitle = stringr::str_wrap(
      "The lower the value, the greater the affinity with Renew's majority. Only the top 5 MEPs by Group are shown.", width = 100) ) +
  theme_minimal() +
  scale_fill_manual(values = polgroup_cols) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "top",
        legend.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"))
# ggsave(here::here("figures", "affinity_renew_meps.jpeg"),
#        dpi=300, device="jpeg", height=3, width=8)
```


#### Share of identical votes
Affinity is defined as the rate with which either the *majorities within Groups/national parties* vote in the same way, or *MEPS* vote in the same way. 
To measure affinity, we consider only EP official votes, that is `for`, `against`, `abstain`.^[Because of ties within Groups (i.e. absence of a majority), we may lose some votes.]

When reading these figures remember that, *contrary to the Euclidean distance*, **high numbers stand for high affinity**, and **low numbers represent low affinity**.
```{r}
#| fig-width: 8
#| fig-height: 4

# plot -------------------------------------------------------------------------
dist_avg_dt <- renew_othermeps[, list(
  same_avg = mean(is_same, na.rm = TRUE) ),
  by = list(pers_id)
]

dist_avg_dt_plot <- dist_avg_dt |>  
  dplyr::inner_join(
    y = meps_current[
      polgroup_id %in% c(7018L, 7038L, 7036L)
    ], by = "pers_id") |> 
  join_polit_labs() |> 
  join_meps_names() |> 
  dplyr::slice_max(order_by = same_avg, n = 3, with_ties = FALSE, 
                   by = political_group) |> 
  dplyr::arrange(political_group, -same_avg)

plot_rank <- dist_avg_dt_plot$mep_name

dist_avg_dt_plot |> 
  ggplot(aes(x = factor(mep_name, levels = plot_rank), 
             y = same_avg, fill = political_group) ) +
  geom_col(colour="black", linewidth=0.1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth=0.5) +
  geom_text(aes(label = round(same_avg*100, digits = 1)), 
            nudge_y = .05, size = 3) +
  facet_grid(cols = vars(political_group), scales = "free_x", space = "free") +
  # scale_x_reordered() +
  scale_x_discrete(labels = scales::wrap_format(10))+
  labs(
    x="", y="Alignment with Renew's Majority (%)", fill="",
    title = "Alignment between Renew's majority and other MEPs",
    subtitle = stringr::str_wrap(
      "The higher the value, the greater the affinity with Renew's majority", 
      width = 100) ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  scale_fill_manual(values = polgroup_cols) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "top",
        legend.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"))
ggsave(here::here("figures", "affinity_renew_meps_overall.pdf"),
       dpi=300, device="pdf", height=4, width=8)
```

```{r}

# plot -------------------------------------------------------------------------
dist_avg_dt <- renew_othermeps[, list(
  same_avg = mean(is_same, na.rm = TRUE) ),
  by = list(pers_id)
]

dist_avg_dt_plot <- dist_avg_dt |>  
  dplyr::inner_join(
    y = meps_current[
      polgroup_id %in% c(7036L)
    ], by = "pers_id") |> 
  join_polit_labs() |> 
  join_meps_names() |> 
  dplyr::slice_max(order_by = same_avg, n = 3, with_ties = FALSE, 
                   by = political_group) |> 
  dplyr::arrange(political_group, -same_avg)

plot_rank <- dist_avg_dt_plot$mep_name

dist_avg_dt_plot |> 
  ggplot(aes(x = factor(mep_name, levels = plot_rank), 
             y = same_avg, fill = mep_name) ) +
  geom_col(colour="black", linewidth=0.1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth=0.5) +
  geom_text(aes(label = round(same_avg*100, digits = 1)), 
            vjust = -0.2, size = 3, fontface = "bold") +
  facet_grid(cols = vars(political_group), scales = "free_x", space = "free") +
  # scale_x_reordered() +
  scale_x_discrete(labels = scales::wrap_format(10))+
  labs(
    x="", y="Alignment with Renew's Majority (%)", fill="",
    title = "Alignment between Renew's majority and other MEPs",
    subtitle = stringr::str_wrap(
      "The higher the value, the greater the affinity with Renew's majority", 
      width = 100) ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "top",
        legend.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"))
ggsave(here::here("figures", "affinity_renew_meps_overall_theleft.jpeg"),
       dpi=300, device="jpeg", height=4, width=8)
```

```{r}

# plot Pappas only May and June 2025 -------------------------------------------
rcvid_tokeep = votes_dt$rcv_id[
  votes_dt$activity_date >= as.Date("2025-05-01") 
  & votes_dt$activity_date <= as.Date("2025-06-30")
  & votes_dt$decision_method == "VOTE_ELECTRONIC_ROLLCALL"
]

dist_avg_dt <- renew_othermeps[
  rcv_id %in% rcvid_tokeep
  & pers_id == 256952L,
  list( same_avg = mean(is_same, na.rm = TRUE) )
]
```


:::


### By Committee

* Between the 9th and 10th mandate, **Elena KOUNTOURA** was a member/Vice-Chair of: Committee on Transport and Tourism (Vice-Chair); Committee on Women’s Rights and Gender Equality; Special Committee on Artificial Intelligence in a Digital Age
* During the 10th mandate, **Nikos PAPPAS** has been a member of: Committee on Industry, Research and Energy; Committee on Culture and Education.

Again, we calculate the affinity metrics in two different ways, as explained above.
As an additional constraint, we require *MEPs to have voted at least 5 times in each Committee* to be included in the metric.

::: {.panel-tabset}

#### Euclidean distance
```{r}
#| fig-width: 8
#| fig-height: 16

# Vector of Committees
cmts_tocheck <- c("AFET", "CULT", "ECON", "ENVI", "FEMM", "IMCO", "ITRE", "LIBE",
                  "TRAN")

rcvid_cmts <- data.table::fread(file = here::here(
  "data_out", "rcv", "rcvid_cmts_all.csv") )
rcvid_cmts_filter <- rcvid_cmts[committee_lab %in% cmts_tocheck]

# Get the target party majority
renew_mjrt_choicecmts <- mjrt_polgroup[
  rcv_id %in% unique(rcvid_cmts_filter$rcv_id)
  & polgroup_id %in% renew_group_ids]

# Inner-join
renew_othermeps_choicecmts <- renew_mjrt_choicecmts[
  meps_rcv_mandate[
    (pers_id %in%  c(197699L, 256952L)
     | polgroup_id %in% c(5153L, 7018L, # EPP
                          5154L, 7038L) ) #S&D
    & rcv_id %in% unique(rcvid_cmts_filter$rcv_id)
    & result >= -1L],
  on = "rcv_id", nomatch = NULL]


# Calculate distances ---------------------------------------------------------#
renew_othermeps_choicecmts[, `:=`(
  is_same = as.integer(result == i.result),
  eucl_dist = sqrt((result - i.result)^2) # 0 if same, 2 if distant
)]


# Merge Affinity with Committee -----------------------------------------------#
renew_othermeps_choicecmts <- renew_othermeps_choicecmts |> 
  dplyr::inner_join(
    y = rcvid_cmts_filter,
    by = "rcv_id") |> 
  data.table::as.data.table()


# Take the Average for Plotting -----------------------------------------------#
dist_avg_dt <- renew_othermeps_choicecmts[, list(
  n_rcv = .N,
  eucl_dist_avg = mean(eucl_dist, na.rm = TRUE) ),
  by = list(pers_id, committee_lab)
][, `:=`(
  # normalise the euclidean for comparable plotting
  eucl_dist_avg_norm = eucl_dist_avg / 2L
)]


# Plot ------------------------------------------------------------------------#
dist_avg_dt[
  n_rcv > 5L,
] |>  
  dplyr::inner_join(
    y = meps_current[
      polgroup_id %in% c(7018L, 7038L, 7036L)
    ], by = "pers_id") |> 
  join_polit_labs() |> 
  join_meps_names() |> 
  dplyr::slice_min(order_by = eucl_dist_avg_norm, n = 5,with_ties = FALSE,
                   by = c(political_group, committee_lab) ) |> 
  dplyr::mutate(mep_name = tidytext::reorder_within(
    x = mep_name, by = eucl_dist_avg_norm, within = committee_lab) ) |> 
  # dplyr::filter(political_group == "PPE") |>
  ggplot(aes(x = mep_name, 
             y = eucl_dist_avg_norm, fill = political_group) ) +
  geom_col(colour="black", linewidth=0.1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth=0.5) +
  facet_wrap(~committee_lab, scales = "free_y", ncol=1) +
  # facet_grid(rows = vars(political_group), scales = "free_y", space = "free") +
  scale_x_reordered() +
  ylim(0, 1) +
  labs(
    x="", y="Normalised euclidean distance", 
    title = "Distance between Renew's majority and other MEPs, by Committee",
    subtitle = stringr::str_wrap(
      "The lower the value, the greater the affinity with Renew's majority. Only the top 5 MEPs by Group are shown. MEPs must have voted at least 5 times in each Committee to be included in the metric.", width = 100) ) +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = polgroup_cols) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "top",
        legend.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"))
```


#### Share of identical votes
```{r}
#| fig-width: 8
#| fig-height: 6

# Take the Average for Plotting -----------------------------------------------#
dist_avg_dt <- renew_othermeps_choicecmts[, list(
  n_rcv = .N,
  same_avg = mean(is_same, na.rm = TRUE) ),
  by = list(pers_id, committee_lab)
]


# Plot ------------------------------------------------------------------------#
dist_avg_dt_plot <- dist_avg_dt[
  !pers_id %in% c(197523, 37324, 86057),
  # & n_rcv > 5L,
] |>  
  dplyr::inner_join(
    y = meps_current[
      polgroup_id %in% c(7018L, 7038L, 7036L)
    ], by = "pers_id") |> 
  join_polit_labs() |> 
  join_meps_names() |> 
  dplyr::slice_max(order_by = same_avg, n = 2, with_ties = FALSE,
                   by = c(political_group, committee_lab) ) |> 
  dplyr::arrange(political_group, -same_avg)

dist_avg_dt_plot |> 
  ggplot(aes(x = mep_name, 
             y = same_avg, fill = political_group) ) +
  geom_col(colour="black", linewidth=0.1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth=0.5) +
  geom_text(aes(label = round(same_avg*100, digits = 1)), 
            nudge_y = .05, size = 3) +
  facet_wrap(~committee_lab, scales = "free_x", ncol=3) +
  # facet_grid(rows = vars(political_group), scales = "free_y", space = "free") +
  scale_x_discrete(labels = scales::wrap_format(5))+
  labs(
    x="", y="Alignment with Renew's Majority (%)", fill="",
    title = "Alignment between Renew's majority and other MEPs, by Committee",
    subtitle = stringr::str_wrap(
      "The higher the value, the greater the affinity with Renew's majority", width = 100) ) +
  theme_minimal() +
  scale_fill_manual(values = polgroup_cols) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1.05)) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "top",
        legend.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text.x = element_text(size = 7, angle = 90, hjust = 1, vjust = 0.5))
ggsave(here::here("figures", "affinity_renew_meps_bycommittee.pdf"),
       dpi=300, device="pdf", height=11, width=8)
```


```{r}
#| fig-width: 8
#| fig-height: 6

# Take the Average for Plotting -----------------------------------------------#
dist_avg_dt <- renew_othermeps_choicecmts[, list(
  n_rcv = .N,
  same_avg = mean(is_same, na.rm = TRUE) ),
  by = list(pers_id, committee_lab)
]


# Plot ------------------------------------------------------------------------#
dist_avg_dt_plot <- dist_avg_dt[
  n_rcv > 5
  & pers_id %in%  c(197699L, 256952L) ] |>  
  dplyr::inner_join(
    y = meps_current, by = "pers_id") |> 
  join_polit_labs() |> 
  join_meps_names() |> 
  dplyr::slice_max(order_by = same_avg, n = 2, with_ties = FALSE,
                   by = c(political_group, committee_lab) ) |> 
  dplyr::arrange(political_group, -same_avg)

dist_avg_dt_plot |> 
  ggplot(aes(x = committee_lab, 
             y = same_avg, fill = mep_name, group = mep_name) ) +
  geom_col(position=position_dodge2(width = 0.9, preserve = "single"),
           colour="black", linewidth=0.1) +
  geom_hline(yintercept = 0, linewidth=0.5) +
  geom_text(aes(label = round(same_avg*100, digits = 1)), #nudge_y = .05, 
            position = position_dodge(0.9), vjust = -0.2, size = 3, fontface = "bold") +
  # facet_wrap(~committee_lab, scales = "free_x", ncol=3) +
  # facet_grid(rows = vars(committee_lab), scales = "free", space = "free") +
  scale_x_discrete(labels = scales::wrap_format(5))+
  labs(
    x="", y="Alignment with Renew's Majority (%)", fill="",
    title = "Alignment between Renew's majority and other MEPs, by Committee",
    subtitle = stringr::str_wrap(
      "The higher the value, the greater the affinity with Renew's majority", width = 100) ) +
  theme_minimal() +
  # scale_fill_manual(values = polgroup_cols) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1.05)) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.position = "top",
        legend.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text.x = element_text(size = 7, face = "bold", hjust = 0.5, vjust = 0.5))
ggsave(here::here("figures", "affinity_renew_meps_bycommittee_theleft.jpeg"),
       dpi=300, device="jpeg", height=4, width=8)
```


```{r}
#| fig-width: 8
#| fig-height: 6

# plot Pappas only May and June 2025 -------------------------------------------
rcvid_tokeep = votes_dt$rcv_id[
  votes_dt$activity_date >= as.Date("2025-05-01") 
  & votes_dt$activity_date <= as.Date("2025-06-30")
  & votes_dt$decision_method == "VOTE_ELECTRONIC_ROLLCALL"
]

# Take the Average for Plotting -----------------------------------------------#
dist_avg_dt <- renew_othermeps_choicecmts[
  rcv_id %in% rcvid_tokeep
  & pers_id == 256952L, 
  list(
    n_rcv = .N,
    same_avg = mean(is_same, na.rm = TRUE) ),
  by = list(committee_lab)
]
```

:::


### Ukraine
```{r}
votes_dt_10 <- data.table::fread(
  file = here::here("data_out", "votes", "votes_dt_10.csv"),
  # select = c("activity_date", "rcv_id", "mandate", "number_of_attendees",
  #            "number_of_votes_abstention", "number_of_votes_against",
  #            "number_of_votes_favor"),
  na.strings = c(NA_character_, "") )

votes_final_10 <- fread(here::here("data_out", "votes", "votes_final_10.csv")) |> 
  dplyr::filter(is_final == 1L)

votes_labels_10 <- fread(here::here("data_out", "votes", "votes_labels_10.csv"))
ukraine_votes = unique(votes_labels_10[
  grepl(pattern = "Ukraine", x = vote_label_en)
  | grepl(pattern = "Ukraine", x = vote_label_fr)
  | grepl(pattern = "Ukraine", x = vote_label_mul)
][["activity_id"]])
votes_dt_10[, vot_id := gsub(pattern = "eli/dl/event/", replacement = "", 
                             x = inverse_consists_of)]
rcvid_ukraine = votes_dt_10[vot_id %in% ukraine_votes][["rcv_id"]]
rcvid_ukraine_finals = rcvid_ukraine[rcvid_ukraine %in% votes_final_10$rcv_id]


dplyr::left_join(
  x = renew_mjrt[rcv_id %in% rcvid_ukraine_finals],
  y = meps_rcv_mandate[
    rcv_id %in% rcvid_ukraine_finals
    & pers_id %in% c(197699L, 256952L),
    list(pers_id, rcv_id, result)],
  by = "rcv_id"
) |> 
  join_meps_names() |> 
  left_join(
    y = votes_dt_10[, list(vot_id, rcv_id)],
    by = "rcv_id"
  ) |> 
  # mutate(vot_id := gsub(pattern = "eli/dl/event/", replacement = "", 
  #                         x = inverse_consists_of)) |> 
  left_join(
    y = unique(votes_labels_10[
      vote_id %in% ukraine_votes, 
      list(vote_label_en, activity_id)]),
    by = c("vot_id" = "activity_id")
  ) |> 
  arrange(rcv_id) |> 
  dplyr::select(
    `Renew Majority` = result.x,
    `MEP Vote` = result.y,
    MEP = mep_name,
    `Vote Label` = vote_label_en
  ) |> 
  fwrite(file = "ukrain.csv")
```



## Data and Methods
```{r}
#| eval: false

# nikos pappas (10)
# elena koutoura (9-10)
# alignement with renew
# commitees (check)
# also check: afet, econ, envi, femm, imco, libe 


#------------------------------------------------------------------------------#
# April PLENARY ----------------------------------------------------------------#
rcvid_target = votes_dt$rcv_id[
  votes_dt$decision_method == "VOTE_ELECTRONIC_ROLLCALL"
  & votes_dt$activity_date %in% 
    seq.Date(from = as.Date("2025-03-31"), 
             to = as.Date("2025-04-03"),
             by = 1)]
meps_rcv_mandate[
  rcv_id %in% rcvid_target
  & pers_id == 256952L,
  list(length(unique(rcv_id))),
  keyby = result
]

merge(x = mjrt_polgroup[polgroup_id %in% renew_group_ids 
                        & rcv_id %in% rcvid_target,
                        list(rcv_id, result)],
      y = meps_rcv_mandate[rcv_id %in% rcvid_target
                           & result >= -1L
                           & pers_id == 256952L],
      by= "rcv_id") |> 
  dplyr::mutate(is_same = as.integer(result.x == result.y) ) |> 
  summarise(avg = mean(is_same, na.rm=TRUE))


# LAST PLENARY ----------------------------------------------------------------#
rcvid_target = votes_dt$rcv_id[
  votes_dt$decision_method == "VOTE_ELECTRONIC_ROLLCALL"
  & votes_dt$activity_date %in% 
    seq.Date(from = as.Date("2025-05-05"), 
             to = as.Date("2025-05-08"),
             by = 1)]
meps_rcv_mandate[
  rcv_id %in% rcvid_target
  & pers_id == 256952L,
  list(unique(rcv_id)),
  keyby = result
]

merge(x = mjrt_polgroup[polgroup_id %in% renew_group_ids 
                        & rcv_id %in% rcvid_target,
                        list(rcv_id, result)],
      y = meps_rcv_mandate[rcv_id %in% rcvid_target
                           & result >= -1L
                           & pers_id == 256952L],
      by= "rcv_id") |> 
  dplyr::mutate(is_same = as.integer(result.x == result.y) ) |> 
  summarise(avg = mean(is_same, na.rm=TRUE))

# ENTIRE 10 ----------------------------------------------------------------#
rcvid_target = votes_dt$rcv_id[
  votes_dt$mandate == 10L]
meps_rcv_mandate[
  rcv_id %in% rcvid_target
  & pers_id == 256952L,
  list(length(unique(rcv_id))),
  keyby = result
]

merge(x = mjrt_polgroup[polgroup_id %in% renew_group_ids 
                        & rcv_id %in% rcvid_target,
                        list(rcv_id, result)],
      y = meps_rcv_mandate[rcv_id %in% rcvid_target
                           & result >= -1L
                           & pers_id == 256952L],
      by= "rcv_id") |> 
  dplyr::mutate(is_same = as.integer(result.x == result.y) ) |> 
  summarise(avg = mean(is_same, na.rm=TRUE))

#------------------------------------------------------------------------------#
```

We include in our analysis all voting data relative to Roll-Call Votes (RCVs) in Plenaries during the current and previous mandates.
To provide some context, we currently have `r length(unique(meps_rcv_mandate$rcv_id))` RCVs in our database for the current mandate.
It is important to appreciate that what we have is only a subset of the all legislation being adopted. 
Indeed, the EP recognises different forms of voting - such as *raise of hands*, or *electronic votes* - but RCVs are the only ones where we able to trace votes to individual MEPs - that is, they are so-called *nominal* votes. 
As a rough approximation, if we take out the likely duplicate rows from the voting archives, RCVs were approximately `r round(n_rcv/n_votes*100, digits=1)`% of the total votes during this mandate.
