---
title: "Communicating about voting data at national level"
author:
  - Marco SCIPIONI
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    toc-title: Contents
    number-sections: true
    colorlinks: true
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  warning: false
---

## Intro
```{r misc}
#| include: false

###--------------------------------------------------------------------------###
## Libraries -------------------------------------------------------------------
if ( !require("pacman") ) install.packages("pacman")
pacman::p_load(char = c("data.table", "dplyr", "future.apply", "ggplot2",
                        "googledrive", "here", "httr", "httr2", "lubridate",
                        "janitor", "jsonlite", "stringi", "tidyr", "tidyselect",
                        "tidytext") )

# Google's credentials
googledrive::drive_auth(email = "marco.scipioni05@gmail.com")


# Hard code the start of the mandate ------------------------------------------#
mandate_starts <- as.Date("2024-07-14")


###--------------------------------------------------------------------------###
## Functions -------------------------------------------------------------------
# https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode
stat_mode <- function(x) {
  if ( length(x) <= 2 ) return(x[1])
  if ( anyNA(x) ) x = x[!is.na(x)]
  ux <- unique(x)
  ux[ which.max( tabulate( match(x, ux) ) ) ] }

# Cohesion rate ---------------------------------------------------------------#
source(file = here::here("source_scripts_r", "cohesionrate_function.R") )

# Join functions --------------------------------------------------------------#
source(file = here::here("source_scripts_r", "join_functions.R") )

# Calculate Majorities --------------------------------------------------------#
source(file = here::here("source_scripts_r", "get_majority.R") )


#------------------------------------------------------------------------------#
## Graphics --------------------------------------------------------------------
# Vote Colours ----------------------------------------------------------------#
vote_colours <- c(For = '#00AEEF',
                  Against = '#BE3455',
                  Abstain = "#969696",
                  `Did not vote` = '#5D5CA4',
                  Absent = '#F47920')

# set theme globally for ggplots ----------------------------------------------#
# REF: https://stackoverflow.com/questions/34522732/changing-fonts-in-ggplot2
theme_set(theme_minimal(base_size = 10, base_family = "roboto"))


###--------------------------------------------------------------------------###
# Country parameters -----------------------------------------------------------
country_name <- "Portugal"
country_iso3 <- "PRT"
country_iso2 <- "PT"
country_int <- country_dict$country_id[country_dict$country == country_iso3]
country_adj <- "Portuguese"
renew_country_party_id <- 6641L
renew_country_party <- national_parties$label[
  national_parties$identifier == renew_country_party_id]
renew_country_party_long <- national_parties$pref_label_en[
  national_parties$identifier == renew_country_party_id]
renew_country_party_noshow_id <- NULL
independents <- c("-", "Ind.", "Indépendant", "Independent", "Independiente")
renew_polgroup_id <- c(5704L, 7035L)
```

The objective of this short note is to provide regular updates to *`r renew_country_party_long`* (from now on, `r renew_country_party`) on Plenary voting dynamics.
At the time of writing, the sample size is extremely low, so the current figures may be misleading. 
Please consider this document just as a template that will become more reliable and insightful as data accumulates. 
That said, if you have specific requests, or you would like to tailor this document according to other specifications, please do not hesitate to get in touch.

**Please keep in mind** that this document is the *sole property of the Renew Group* and *should not be disseminated to other people or organisations*.


## Data and Methods
We include in our analysis all voting data relative to RCVs since the start of the 10th mandate.
This means that we also get several MEPs who switched party and/or Group.
All these changes may translate into MEPs and parties featuring in different EP Political Groups.
For the sake of simplicity, in this analysis we discard MEPs' vote intentions, and just display calculations relative to `r country_adj` MEPs and parties.
```{r}
#| include: false

###--------------------------------------------------------------------------###
## Read data -------------------------------------------------------------------

if (file.exists( here::here("data_out", "meps_rcv_mandate_10.csv") ) ) {
  # RCVs ----------------------------------------------------------------------#
  meps_rcv_mandate <- data.table::fread(
    file = here::here("data_out", "meps_rcv_mandate_10.csv"),
    verbose = TRUE, key = c("rcv_id", "pers_id"), 
    na.strings = c(NA_character_, "")  )
  # Votes ---------------------------------------------------------------------#
  votes_dt <- data.table::fread(
    file = here::here("data_out", "votes", "votes_dt_10.csv"), 
    select = c("activity_date", "rcv_id", "mandate", "decision_method",
               "number_of_attendees", "number_of_votes_abstention", 
               "number_of_votes_against", "number_of_votes_favor"), 
    na.strings = c(NA_character_, "")  )  
} else {
  # Download via Google Drive -------------------------------------------------#
  ## have you gone through authentication? ------------------------------------#
  data_id <- googledrive::drive_get("data_out/meps_rcv_mandate_10.csv")
  ## download -----------------------------------------------------------------#
  googledrive::drive_download(file = data_id, path = here::here(
    "data_out", "meps_rcv_mandate_10.csv"), overwrite = T)
  meps_rcv_mandate <- data.table::fread(
    file = here::here("data_out", "meps_rcv_mandate_10.csv"),
    verbose = TRUE, key = c("rcv_id", "pers_id"), 
    na.strings = c(NA_character_, "")  ) 
  
  ### Votes --------------------------------------------------------------------
  votes_dt <- data.table::fread(
    file = here::here("data_out", "votes", "votes_dt_10.csv"), 
    select = c("activity_date", "rcv_id", "mandate", "decision_method",
               "number_of_attendees", "number_of_votes_abstention", 
               "number_of_votes_against", "number_of_votes_favor"), 
    na.strings = c(NA_character_, "")  )  
}

# Vote dictionary -------------------------------------------------------------#
vote_dict <- data.table::fread(here::here("data_out", "votes", "vote_dict.csv") )


### MEP last Plenary day -------------------------------------------------------
meps_current <- data.table::fread( file = here::here(
  "data_out", "meps", "meps_current.csv"), 
  na.strings = c(NA_character_, "")  )

# get country nat_party id
country_parties_current <- sort( unique( meps_current$natparty_id[
  meps_current$country_id == country_int] ) ) 

# full list of Renew MEPs
renew_meps_full <- unique(meps_rcv_mandate$pers_id[
  meps_rcv_mandate$polgroup_id %in% renew_polgroup_id])


###--------------------------------------------------------------------------###
### MEPs mandate ---------------------------------------------------------------
meps_mandate <- data.table::fread(
  here::here("data_out", "meps", "meps_mandate_10.csv"))
```


## Best people
```{r}
# Get the Groups' majority by rcv_id ------------------------------------------#
who_won_bygroup <- get_polgroup_majority(
  data_in = meps_rcv_mandate[result >= -2]) # exclude absent

# Get Party majority ----------------------------------------------------------#
who_won_byparty <- get_natparty_majority(
  data_in = meps_rcv_mandate[result >= -2]) # exclude absent

## Calculate similarity --------------------------------------------------------
# Merge Group line with MEPs vote 
who_won_group_mep <- who_won_bygroup[
  meps_rcv_mandate, on = c("rcv_id", "polgroup_id") 
][i.result >= -2] # exclude absent

# Are Groups' majorities and MEP the same? ------------------------------------#
who_won_group_mep[, is_same := as.integer(result == i.result)]
```
### Best people to talk to per session
What MEPs (not from Renew) voted in line with `r renew_country_party` when they deviated from their own group line?
Here we consider all types of votes, including the `no_vote` (i.e. a MEP present in the House but not voting), essentially getting rid only of those who are absent.
The vote dictionary is as follows:
```{r}
vote_dict |> 
  dplyr::filter(result >= -2) |> 
  knitr::kable(align = "c")
```

To answer that question, we first have to check whether non-Renew MEPs deviates from their Group.
The bar chart below displays how many times MEPs have deviated from the Groups' majorities during the last Plenary.
```{r}
#| fig-width: 8
#| fig-height: 3

#------------------------------------------------------------------------------#
# First condition: non-Renew parties deviates from their Group
votes_dt[, `:=`(
  week_n = data.table::week(activity_date),
  year_n = data.table::year(activity_date)
)]
# Check 
# votes_dt[decision_method == "def/ep-decision-methods/VOTE_ELECTRONIC_ROLLCALL",
#              length(unique(rcv_id)),
#              by = activity_date]

# Extract vector of RCV IDs from last Plenary
rcvids_lastplenary <- unique(votes_dt$rcv_id[
  votes_dt$decision_method == "def/ep-decision-methods/VOTE_ELECTRONIC_ROLLCALL" # RCV only
  & votes_dt$week_n == max(votes_dt$week_n) # last Plenary only
  & votes_dt$year_n == max(votes_dt$year_n) ] )

# Filter on just MEPs' deviations from Groups ---------------------------------#
group_meps_deviations <- who_won_group_mep[
  !polgroup_id %in% renew_polgroup_id # no Renew
  & is_same == 0L] |>  # mismatches only 
  join_polit_labs() |> 
  join_meps_names()

# Aggregate & plot ------------------------------------------------------------#
group_meps_deviations[
  country_id == country_int
  & rcv_id %in% rcvids_lastplenary, # subset to last Plenary
  list(Count = .N), keyby = list(political_group) ] |> 
  dplyr::mutate(political_group = forcats::fct_reorder(political_group, Count)) |> 
  ggplot(aes(x = political_group, y = Count)) +
  geom_col(linewidth = 0.1, colour = "black", fill = "grey80") +
  # geom_text(aes(label = Count), nudge_y = 0.5, size = 3) +
  labs(x = "", y = "Count", fill = "",
       title = "Count of MEPs' deviations from Group's majority - Last Plenary",
       caption = "Note: data refers only to RCVs, other types of EP votes are not included.") +
  theme_minimal() +
  coord_flip() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.text = element_text(face="bold"),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"),
        legend.position = "none")
```

Having isolated the MEPs who deviated from the the Groups' majorities, we then proceed to single out whether any of those voted alongside `r renew_country_party`.
The table below lists the top 20 MEPs who voted more frequently in the same way as `r renew_country_party` and who did not follow their own Groups' majorities.
```{r}
# Merge Renew Party's majority with deviant MEPS ------------------------------#
renewparty_deviantmeps <- merge(
  x = who_won_byparty[
    rcv_id %in% rcvids_lastplenary # subset to last Plenary
    & natparty_id %in% renew_country_party_id, # Just target Party
    list(rcv_id, result)],
  y = group_meps_deviations[
    country_id == country_int
    & rcv_id %in% rcvids_lastplenary, # subset to last Plenary
    list(rcv_id, i.result, country_id, polgroup_id, political_group, natparty_id,
         national_party, pers_id, mep_name) ],
  by = "rcv_id", all = FALSE)

# Are Groups' majorities and MEP the same? ------------------------------------#
renewparty_deviantmeps[, is_same := as.integer(result == i.result)]

# Aggregate and print table ---------------------------------------------------#
renewparty_deviantmeps[, list(
  Count = sum(is_same, na.rm = TRUE)), 
  keyby = list(country_id, political_group, national_party, mep_name) ] |>
  join_meps_countries() |> 
  dplyr::select(country, political_group, national_party, mep_name, Count) |> 
  dplyr::arrange(desc(Count)) |> 
  head(20) |> 
  knitr::kable(align = "c")
```


### Best people to talk to in the parliamentary term
In this Section, we basically repeat the same calculations as above, only this time without any filtering to just the last Plenary. 
Thus, as above, the bar chart below displays how many times MEPs have deviated from the Groups' majorities since the start of the mandate. 
```{r}
#| fig-width: 8
#| fig-height: 3

#------------------------------------------------------------------------------#
# Aggregate & plot ------------------------------------------------------------#
group_meps_deviations[
  country_id == country_int,
  list(Count = .N),
  keyby = list(political_group) ] |> 
  dplyr::mutate(political_group = forcats::fct_reorder(political_group, Count)) |> 
  ggplot(aes(x = political_group, y = Count)) +
  geom_col(linewidth = 0.1, colour = "black", fill = "grey80") +
  # geom_text(aes(label = Count), nudge_y = 3, size = 3) +
  labs(x = "", y = "Count", fill = "",
       title = "Count of MEPs' deviations from Group's majority - Full mandate",
       caption = "Note: data refers only to RCVs, other types of EP votes are not included.") +
  theme_minimal() +
  coord_flip() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.text = element_text(face="bold"),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"),
        legend.position = "none")
```

Having isolated the MEPs who deviated from the the Groups' majorities, we then proceed to single out whether any of those voted alongside `r renew_country_party`.
The table below lists the top 20 MEPs who voted more frequently in the same way as `r renew_country_party` and who did not follow their own Groups' majorities.
```{r}
# Merge Renew Party's majority with deviant MEPS ------------------------------#
renewparty_deviantmeps <- merge(
  x = who_won_byparty[
    natparty_id %in% renew_country_party_id, # Just target Party
    list(rcv_id, result)],
  y = group_meps_deviations[
    country_id == country_int,
    list(rcv_id, i.result, country_id, polgroup_id, political_group, natparty_id,
         national_party, pers_id, mep_name) ],
  by = "rcv_id", all = FALSE)

# Are Groups' majorities and MEP the same? ------------------------------------#
renewparty_deviantmeps[, is_same := as.integer(result == i.result)]

# Aggregate and print table ---------------------------------------------------#
renewparty_deviantmeps[, list(
  Count = sum(is_same, na.rm = TRUE)), 
  keyby = list(country_id, political_group, national_party, mep_name) ] |>
  join_meps_countries() |> 
  dplyr::select(country, political_group, national_party, mep_name, Count) |> 
  dplyr::arrange(desc(Count)) |> 
  head(20) |> 
  knitr::kable(align = "c")
```

If we focus just on EPP, the top 5 MEPs are:
```{r}
# Aggregate and print table ---------------------------------------------------#
renewparty_deviantmeps[, list(
  Count = sum(is_same, na.rm = TRUE) ),
  keyby = list(country_id, political_group, national_party, mep_name) ] |>
  join_meps_countries() |> 
  dplyr::select(country, political_group, national_party, mep_name, Count) |> 
  dplyr::arrange(desc(Count)) |> 
  dplyr::filter(political_group == "PPE") |> 
  head(5) |> 
  knitr::kable(align = "c")
```


### Best people to talk to in the parliamentary term by Committee
**NOT CURRENTLY RUNNING THIS AS WE DON'T HAVE ENOUGH DATA**
```{r}
#| include: false
#| fig-width: 8
#| fig-height: 10 

#------------------------------------------------------------------------------#
# List of RCV IDs and Doc IDs -------------------------------------------------#
rcvid_docid_all <- data.table::fread( here::here(
  "data_out", "rcv", "rcvid_docid_10.csv"), 
  na.strings = c(NA_character_, "")  ) 

# Committees -------------------------------------------------------------------
plenary_docs_committee <- data.table::fread( here::here(
  "data_out", "docs_pl", "plenary_docs_committee.csv"), 
  na.strings = c(NA_character_, "")  ) |> 
  dplyr::select(-identifier)

# Merge RCV ID, Plenary Docs, and Committees ----------------------------------#
pldoc_cmmtt_rcvid <- plenary_docs_committee[
  rcvid_docid_all, 
  on = "doc_id", 
  nomatch = NULL] # inner join

target_committees <- c("CONT", "CULT", "ECON", "EMPL", "ENVI", "FEMM", "ITRE",
                       "LIBE", "REGI")

# Merge Plenary Docs, RCV, Committees with deviant MEPs -----------------------#
dt_plot <- renewparty_deviantmeps[
  is_same == 1L # filter to just same votes
][
  pldoc_cmmtt_rcvid,
  on = "rcv_id", 
  nomatch = NULL] # inner join

# Aggregate and plot ----------------------------------------------------------#
dt_plot[, list(
  Count = sum(is_same, na.rm = TRUE) 
), 
keyby = list(country_id, political_group, national_party, mep_name,
             committee_lab) ] |>
  join_meps_countries() |> 
  # filter to just selected Committees ----------------------------------------#
  dplyr::filter(committee_lab %in% target_committees) |> 
  dplyr::mutate(mep_name = tidytext::reorder_within(
    mep_name, Count, committee_lab) ) |> 
  ggplot(aes(x = mep_name, y = Count)) +
  geom_col(linewidth = 0.1, colour = "black", fill = "grey80") +
  # geom_text(aes(label = Count), nudge_y = 3, size = 3) +
  facet_wrap(~committee_lab, scales = "free_y") +
  labs(x = "", y = "Count", fill = "",
       title = "Count of MEPs' deviations from Group's majority - Full mandate",
       caption = "Note: data refers only to RCVs, other types of EP votes are not included.") +
  coord_flip() +
  tidytext::scale_x_reordered() +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.text = element_text(face="bold"),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"),
        legend.position = "none")
```


## Opposition
In this section we focus on `r renew_country_party`'s main opposition parties in the House. 


### Opposition parties' divergence with their respective Groups during the last Plenary
In this Section, we select the instances where `r country_adj` parties did not vote together with their own respective Political Groups in the last Plenary.
We toss out **independents** to simplify the presentation.
They can be brought back in at any time if needed. 
```{r}
who_won_group_party <- unique(
  # get the unique combination to anchor joins
  meps_rcv_mandate[
    country_id == country_int
    & !natparty_id %in% natparties_independent_ids, # no INDEPENDENT
    list(rcv_id, country_id, polgroup_id, natparty_id) ] ) |> 
  # inner join with Political Groups majority
  dplyr::inner_join(
    y = who_won_bygroup, by = c("rcv_id", "polgroup_id") ) |> 
  # inner join with National Parties majority
  dplyr::inner_join(
    y = who_won_byparty, by = c("rcv_id", "natparty_id") ) |> 
  data.table::as.data.table()


# Are Group and party majorities the same? ----------------------------------###
who_won_group_party[, is_same := as.integer(result.x == result.y) ]

who_won_group_party[
  country_id == country_int # get just data from the target country
  & !polgroup_id %in% renew_polgroup_id
  & is_same == 0L
  & rcv_id %in% rcvids_lastplenary # filter to just last Plenary
] |> 
  join_polit_labs() |> 
  dplyr::select(rcv_id, political_group, national_party,
                `Group Majority` = result.x,
                `Party Majority` = result.y) |> 
  dplyr::arrange(rcv_id, political_group, national_party) |> 
  knitr::kable(align = "c")
```


### Opposition parties' alignment with their respective Groups since the start of the mandate
Now, instead of just focusing on the last Plenary, we calculate the frequency with which `r country_adj` parties have voted together with their own Groups since the start of the mandate.
In this Section, we toss out **independents** to simplify the presentation.
They can be brought back in at any time if needed. 
```{r}
#| fig-width: 8
#| fig-height: 5

who_won_group_party_avg <- who_won_group_party[
  country_id == country_int # get just data from the target country
  & !polgroup_id %in% renew_polgroup_id,
  list(
    align_share = mean(is_same, na.rm = TRUE) ),
  by = list(polgroup_id, natparty_id)
] 

who_won_group_party_avg |> 
  join_polit_labs() |> 
  dplyr::mutate(national_party = forcats::fct_reorder(
    national_party, align_share)) |> 
  ggplot(aes(x = national_party, y = align_share)) +
  geom_col(linewidth = 0.1, colour = "black", fill = "grey80") +
  # geom_text(aes(label = Count), nudge_y = 3, size = 3) +
  facet_grid(rows = "political_group", scales = "free", space = "free") +
  labs(x = "", y = "Alignment Share (%)", fill = "",
       title = "National Parties' alignment with the respective Group's majority - Full mandate",
       caption = "Note: data refers only to RCVs, other types of EP votes are not included.") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  coord_flip() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.text = element_text(face="bold"),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0, face="bold"),
        legend.position = "none")
```


### Opposition parties' internal cohesion rates since the start of the mandate
Having looked at parties alignment with their respective Groups, we now check the extent to which `r country_adj` parties are internally cohesive. 
```{r}
#| fig-width: 8
#| fig-height: 3

#------------------------------------------------------------------------------#
# left join
meps_rcv_mandate <- vote_dict[meps_rcv_mandate, on = "result"]
# convert to factor - essential for tabulate
meps_rcv_mandate[, result_fct := factor(result_fct,
                                        levels = c("absent", "no_vote", "against",
                                                   "abstain", "for") ) ]
# Calculate Cohesion
cohesion_dt <- meps_rcv_mandate[
  country_id == country_int
  & result >= -1L, # only official votes
  list(
    cohesion = cohesion_hn(result_fct)
  ), 
  keyby = list(rcv_id, polgroup_id, natparty_id) ] 

cohesion_avg <- cohesion_dt[
  ! is.na(cohesion)
  & ! polgroup_id %in% renew_polgroup_id,
  list(
    cohesion_avg = mean(cohesion, na.rm = TRUE) ),
  by = list(polgroup_id, natparty_id)
] 

cohesion_avg |> 
  join_polit_labs() |> 
  dplyr::mutate(national_party = forcats::fct_reorder(
    national_party, cohesion_avg)) |> 
  ggplot(aes(x = national_party, y = cohesion_avg)) +
  geom_col(linewidth = 0.1, colour = "black", fill = "grey80") +
  facet_grid(rows = "political_group", scales = "free", space = "free") +
  labs(x = "", y = "Cohesion rate (%)", fill = "",
       title = "National Parties' internal cohesion rates - Full mandate",
       caption = "Note: data refers only to RCVs, other types of EP votes are not included.") +
  theme_minimal() +
  coord_flip() +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        legend.text = element_text(face="bold"),
        panel.grid.major.y = element_line(colour = "grey30", linewidth = 0.1),
        panel.grid.minor.y = element_line(colour = "grey70"),
        axis.text = element_text(face="bold"),
        strip.text.y = element_text(angle = 0, face="bold"),
        legend.position = "none")
```
